---
title: "Assignment 4"
output: html_notebook
---

ECE 608 (Spring 2019) Assignment 4
Professor: Jason Au
Student: Juan Manuel Gomez Gonzalez


Information: 
Power and Sample Size Calculations 

First, the required libraries need to be loaded:
```{r Libraries load}
library(tidyverse)
library(grid)
library(gridExtra)
library(pwr)
library(pwr2)
```

# Question 1
## a.
The purpose of the study is to review both published and unpublished literature, in order to find if the process of active learning helps students achieve higher learning scores in STEM related undergraduate degrees, compared with traditional learning. They mainly analyzed scores in identical or equivalent testing schemes and failure rates (D or F marks) in the analyzed courses.

## b.
Fig 2.A. shows a Forest Plot for the effect sizes divided by each discipline. It basically shows the effect size of the comparison between active learning and traditional lecturing in 8 the STEM fields analyzed.

## c.
One of the primary findings is that when talking about examination results or equivalent measurements, there is a statistically significant increase in performance (with a weighted standardized mean difference of approximately 0.5 deviations, SD (Z = 9.781, P << 0.001)) when using active learning in comparison with only traditional learning. The other primary finding is that there is an odds ratio of 1.95 (Z = 10.4, P << 0.001) for the overall mean effect size. The researchers mention that this 1.95 odds ratio is equivalent to a risk ratio of 1.5, where there is a 1.5 times bigger risk of failing a course when using traditional lecturing compared with active learning.

## d.
The main difference seems to come from the way each statistic calculates the pooled standard deviation. Cohen's uses the pooled standard deviation for the dividend of the formula, while Hedge's uses the weighted standard deviation [1].


This makes the Hedge's preferable, as it has been found that it has better properties when the sample sizes are significantly different or when dealing with smaller sample sizes [2], specifically less than 20 [1]. Nonetheless, their performance is similar when dealing with large samples where both sample sizes are similar [2].

According to this, it seems that the choice for using Hedge's instead of Cohen's is mainly due to the size of some of the samples of the papers, where they are found to be as small as n=16.

# Question 2
## a.
Using the overall effect size found in the research paper of 0.47 SD, and assuming a significance level of 0.01 and a power of 0.8, it is possible to calculate the number of samples needed to obtain the same effect present on the paper:

```{r Q2.a Sample size calculation}
result <- pwr.t.test(d = 0.47, #effect size
           n = NULL, #Sample size
           sig.level = 0.01, #alpha
           power = 0.8, #power
           type = "two.sample") #Specific for 2 sample  t-test
print(result)
writeLines(paste("Total sample size is", 2 * ceiling(result[["n"]])))
```

A minimum of 108 samples are needed per group, for a total of 216, to obtain an effect size of 0.47 SD when the significance level is 0.01 and the power is 0.8.

## b.
Assuming a significance level of 0.01 and a power 0f 0.8, the ss.2way function can be used to calculate the sample sizes needed when the expected f-score is 0.5 for learning style and 0.2 for STEM discipline:

```{r Q2.b Sample size calculation for 2way ANOVA}
ss.2way(a = 2, #learning style
        b = 8, #STEM disciplines
        alpha = 0.01, #significance level
        beta = 0.20, #beta = (1 - power)
        f.A = 0.5, #f score for a
        f.B = 0.2, #f score for b
        B=1000) #iteration times
```
32 samples are needed per group, for a total minimum of 512 samples for obtaining an f value of 0.5 for learning style and 0.2 for STEM discipline guaranteeing the significance and power levels previously selected.

## c.
To correctly create the different plots, we need to sequentially calculate the power through different sample sizes for the effect size of each STEM discipline.
```{r Q2.c Power vs Sample Size Plots, fig.width=10, fig.height = 6}
# Biology = 0.303
# Engineering = 0.481
# Math = 0.340

n <- seq(5, 300, by = 1)
lenn <- length(n)
d <- c(0.303, 0.481, 0.340)
lend <- length(d)

pwrtable <- array(numeric(lenn*lend), dim = c(lenn,lend))
for (i in 1:lend){
  for (j in 1:lenn){
    result <- pwr.t.test(d = d[i],
                         n = n[j],
                         sig.level = 0.01,
                         type = "two.sample",
                         alternative = "two.sided")
    pwrtable[j,i] <- result$power
  }
}

pwrtable <- data.frame(n, pwrtable)
names(pwrtable) <- c("samplesize","power_Bio","power_Eng","power_Math")

#Create the Biology plot
bio_plot <- ggplot(data = pwrtable, aes(x = samplesize, y = power_Bio)) +
  labs(y = 'Power (Biology)', x = 'Sample Size') +
  ylim(0,1) +
  theme_light() +
  geom_line(color="forestgreen", aes(color = "Biology")) +
  geom_hline(yintercept = 0.80) +
  geom_point(data = data.frame(samplesize = pwrtable[match(TRUE, pwrtable[["power_Bio"]]>0.8),"samplesize"], power_Bio = 0.8)) +
  geom_text(data = data.frame(samplesize = 0.85 * pwrtable[match(TRUE, pwrtable[["power_Bio"]]>0.8),"samplesize"], power_Bio = 0.82),
                              label = paste("n =", pwrtable[match(TRUE, pwrtable[["power_Bio"]]>0.8),"samplesize"]), size = 4)

#Create the Engineering plot
eng_plot <- ggplot(data = pwrtable, aes(x = samplesize, y = power_Eng)) +
  labs(y = 'Power (Engineering)', x = 'Sample Size') +
  ylim(0,1) +
  theme_light() +
  geom_line(color="Purple", aes(color = "Engineering")) +
  geom_hline(yintercept = 0.80) +
  geom_point(data = data.frame(samplesize = pwrtable[match(TRUE, pwrtable[["power_Eng"]]>0.8),"samplesize"], power_Eng = 0.8)) +
  geom_text(data = data.frame(samplesize = 0.85 * pwrtable[match(TRUE, pwrtable[["power_Eng"]]>0.8),"samplesize"], power_Eng = 0.82),
                              label = paste("n =", pwrtable[match(TRUE, pwrtable[["power_Eng"]]>0.8),"samplesize"]), size = 4)

#Create the Math plot
math_plot <- ggplot(data = pwrtable, aes(x = samplesize, y = power_Math)) +
  labs(y = 'Power (Math)', x = 'Sample Size') +
  ylim(0,1) +
  theme_light() +
  geom_line(color="gold", aes(color = "Math")) +
  geom_hline(yintercept = 0.80) +
  geom_point(data = data.frame(samplesize = pwrtable[match(TRUE, pwrtable[["power_Math"]]>0.8),"samplesize"], power_Math = 0.8)) +
  geom_text(data = data.frame(samplesize = 0.85 * pwrtable[match(TRUE, pwrtable[["power_Math"]]>0.8),"samplesize"], power_Math = 0.82),
                              label = paste("n =", pwrtable[match(TRUE, pwrtable[["power_Math"]]>0.8),"samplesize"]), size = 4)


# Create the caption. This part is based on an answer given by hrbmstr in StackOverflow: https://stackoverflow.com/a/41106250
figCaption <- textGrob(paste("Figure 1. Power vs Sample Size plots for the departments of Biology (left), Engineering (middle), and Math (right).",
                       "\nThe number after the n represents the sample size needed to achieve a power value equal or greater than 0.8"), 
           gp = gpar(fontsize=9, col = "#2b2b2b"), 
           x = unit(0.005, "npc"), just = c("left", "bottom"))

# Build the layout
plot_layout <- rbind(c(1,2,3), c(4,4,4))

# Plot everything
grid.arrange(bio_plot, eng_plot, math_plot, figCaption, layout_matrix = plot_layout, heights = c(0.9, 0.1))
```
 
The most appropriate size that would guarantee an effect size with a power of 0.8 and a significance level of 0.01 would be 257 samples for each of the STEM disciplines considered in this exercise.
 
# Question 3
Results from Freeman et al (2014) [3] were used to estimate the effect size for finding a change in performance on examinations or equivalent tests (overall effect size), with an effect size of d = 0.47 SD, alpha of 0.01 and power of 0.8 yielded a sample size of 108 per group. Anticipating problems occuring from the possibility of not correctly applying the active learning methods on the student population, we aim to study 140 students per group.

# References
[1] 	Statistics How To, "Hedges'g: Definition, Formula," Statistics How To, 2019. [Online]. Available: https://www.statisticshowto.datasciencecentral.com/hedges-g/. [Accessed 1 July 2019].

[2] 	National Institute of Standards and Technology - NIST, "Hedge's g Statistic," National Institute of Standards and Technology, 15 August 2018. [Online]. Available: https://www.itl.nist.gov/div898/software/dataplot/refman2/auxillar/hedgeg.htm. [Accessed 1 July 2019].

[3] 	S. Freeman, S. L. Eddy, M. McDonough, M. K. Smith, N. Okorafor, H. Jordt and M. P. Wenderoth, "Active learning increases student performance in science, engineering, and mathematics," Proceedings of the National Academy of Sciences of the United States of America (PNAS), vol. 111, no. 23, pp. 8410 - 8415, 2014. 