---
title: "Assignment 6"
output: html_notebook
---

ECE 608 (Spring 2019) Assignment 6
Professor: Jason Au
Student: Juan Manuel Gomez Gonzalez


Information: 
Reliability and Validity


First, the required libraries need to be loaded:
```{r Libraries load, results='hide'}
library(tidyverse)
library(psych)
library(BlandAltmanLeh)
```

# Question 1
Gus T. Wind is checking the reliability of the machine itself, and will only use this machine for his tests. This makes the model of the reliability test a Two-Way Mixed as it is an Test-Retest reliability test. Additionaly, the idea of the test is to obtain the true values of the oxygen and CO2 concentrations, therefore the test is of the absolute agreement type.
Finally, Gus T. Wind intents on only using a signle measurement for the machine, so it is a single measures instead of an average measures test.
The test can thus be classified as an ICC(3, 1) Absolute Agreement, Single Measures.
The reliability score can then be calculated using the data find in the dataframe, which must be loaded first.
```{r DataFrame load}
df <- data.table::fread("A6_Reliability.csv")
```

Afterwards, the ICC function can be employed to obtain the different parameters. The value of interest ins then the ICC3 type found in the return of this function.
```{r Q1 ICC}
ICC_1 <- psych::ICC(df %>% select(Gas1, Gas2, Gas3, Gas4, Gas5, Gas6))
ICC_1[["results"]] %>% 
  tibble::rownames_to_column() %>% 
  dplyr::filter(str_detect(rowname, 'Single_fixed_raters'))
```

The ICC for the gas analyzer is 0.76.

# Question 2
Cy Kyle Fast and company are trying to look at inter-rater reliability, with the condition that they want to make it as generalizable as possible. This means that for this case, it is necessary to use a Two-Way Random model. They are trying to measure the con

1

 231`212121
 
   sistency of the measurements, and so they should use the consistency type. Finally, they mention that 
Due to the previous statements, the test can be classified as an ICC(2, 6) 

```{r Q2 ICC}
ICC_1 <- psych::ICC(df %>% select(Tired1, Tired2, Tired3, Tired4, Tired5, Tired6))
ICC_1[["results"]] %>% tibble::rownames_to_column() %>% dplyr::filter(str_detect(rowname, 'Single_fixed_raters'))
```

# Question 3
Juan R. Em wants to check his own reliability, meaning that we're looking at intra-rater reliability. This means that we are to use a two-way mixed test.